---
title: "The dataset.yaml spec"
---

Every dataset has a spec file at `harmonisation/datasets/BPIPD-<id>/dataset.yaml` that tells the pipeline what files exist, where they live, and how to read them.
This page is the field-by-field reference; for worked examples of common layouts, see the [Spec cookbook](spec-cookbook.qmd).

The shape of the file is defined by `harmonisation/dataset.schema.json`.
A minimal spec looks like this:

```yaml
dataset_id: "9999"
dataset_name: "Example Cohort Study"
status: "in_progress"

resources:
  - name: "data"
    role: "data"
    glob: "data.sav"
    reader: "spss"
```

## Top-level fields

| Field | Required? | Notes |
|-------|-----------|-------|
| `dataset_id` | Yes | The dataset number **only**, as a string of digits (e.g. `"9999"`, not `"BPIPD-9999"`). Must match the folder name. |
| `dataset_name` | Yes | The name shown next to the dataset ID on the original data folder, e.g. `"Example Cohort Study"`. |
| `status` | Yes | One of `planned`, `in_progress`, `harmonised`, `blocked`. Leave as `in_progress` until harmonisation is complete. Only `in_progress` and `harmonised` datasets are included in the pipeline. |
| `resources` | Optional | File list for datasets **without waves** (see below). |
| `waves` | Optional | File lists per wave for datasets **with waves** (see below). |
| `notes` | Optional | Free-text notes about the dataset. |

A dataset normally has *either* `resources` *or* `waves` — not both, unless there are dataset-wide files that sit outside any single wave (such as a shared codebook), in which case a top-level `resources` list can coexist with `waves`.

## Resources

Each item in `resources` describes one file — or one group of related files — in the dataset.

| Field | Required? | Notes |
|-------|-----------|-------|
| `name` | Yes | A short label, e.g. `"data"`, `"codebook"`. Names must be unique across the whole file (including across waves) — this is the name the tidier uses to access the table. |
| `role` | Yes | One of `data`, `codebook`, `docs` (see below). |
| `glob` | Yes | The filename, including the folder path if nested (see below). |
| `reader` | **Only for `role: data`** | How to read the file (see the reader table). |
| `sheet` | Excel only | Sheet name or 1-based index. |
| `range` | Excel only, optional | A cell range, e.g. `"A1:H200"`. |
| `table` | Access (`mdb`) only | The table to read; may be omitted when the file contains exactly one table. |
| `object` | `rda` only | The object to extract; may be omitted when the file contains exactly one object. |

### Roles

| Role | Use for |
|------|---------|
| `data` | The actual dataset file(s) — `.sav`, `.dta`, `.csv`, `.xlsx`, etc. **Requires a `reader`.** Read into R and passed to the tidier. |
| `codebook` | Data dictionaries and variable documentation. The file path is recorded for reference, but the file is **not read into R**. |
| `docs` | Anything else: technical reports, user guides, questionnaires, syntax files. Paths only — must not have a `reader`, `sheet`, or `range`. |

### Globs

`glob` locates files relative to the dataset's root folder (or the wave's `wave_dir`, for wave resources).

- **File in the dataset folder**: `glob: "data.sav"`
- **File nested in a subfolder** — include the full path: `glob: "Survey Files/2000/data.dta"`
- **Wildcards** — `*` and `?` are supported and can match several files: `glob: "stata/*.dta"`.
  Each match becomes its own table, named `<name>__1`, `<name>__2`, … (see [Tidy scripts](tidy-scripts.qmd#what-the-tidier-receives)).
- **Several named files in one resource** — as a YAML list:

  ```yaml
  glob:
    - "file1.sav"
    - "file2.sav"
  ```

  or with brace notation: `glob: "{file1.sav,file2.sav}"`.
  Both are equivalent; pick whichever reads better and stay consistent within a file.

Paths must match the real files exactly, including capitalisation and spacing — typos in globs are the most common reason a spec reads nothing.

## Readers

`reader` is required whenever `role: data`.
Pick it from the file extension:

```{r}
#| label: reader-table
readers <- tibble::tribble(
  ~Reader, ~`File types`, ~`Extra keys`,
  "csv", "`.csv` (comma-separated)", "",
  "csv2", "`.csv` (semicolon-separated, comma decimal)", "",
  "tsv", "`.tsv` (tab-separated)", "",
  "stata", "`.dta` (Stata)", "",
  "spss", "`.sav` (SPSS)", "",
  "sas", "`.sas7bdat` (SAS)", "",
  "rds", "`.rds` (single R object)", "",
  "rda", "`.rda` / `.RData` (saved R workspace)",
  "`object` (optional if the file holds one object)",
  "parquet", "`.parquet` (Apache Parquet)", "",
  "excel", "`.xlsx` / `.xls` (Excel)",
  "`sheet` (recommended), `range` (optional)",
  "mdb", "`.mdb` / `.accdb` (Microsoft Access)",
  "`table` (optional if the file holds one table)"
)
knitr::kable(readers)
```

```{r}
#| label: reader-drift-check
#| include: false
# Fail the site build if the table above drifts from the schema's enum.
schema <- jsonlite::read_json("../harmonisation/dataset.schema.json")
schema_readers <- unlist(
  schema$definitions$resource$properties$reader$enum
)
stopifnot(setequal(readers$Reader, schema_readers))
```

For **Excel data files**, always set `sheet` — open the file in the original data folder to check the exact sheet name:

```yaml
- name: "questionnaire"
  role: "data"
  glob: "questionnaire.xlsx"
  reader: "excel"
  sheet: "Data"
```

(`reader` and `sheet` are only meaningful on `role: data` resources — codebooks are recorded by path but never read.)

::: {.callout-warning}
`.rda` files use `reader: "rda"`, not `rds`.
An `.rds` file stores exactly one object; an `.rda` file is a saved workspace that can hold several, which is why `rda` takes an optional `object` key.
:::

## Waves

If a dataset has files for multiple waves (years, sweeps), use `waves` instead of a top-level `resources` list:

```yaml
waves:
  - wave: "2010"
    label: "2010"
    wave_dir: "."
    resources:
      - name: "data_2010"
        role: "data"
        glob: "survey_2010.csv"
        reader: "csv"
```

| Field | Required? | Notes |
|-------|-----------|-------|
| `wave` | Yes | The wave identifier, e.g. `"2001"`, `"sweep_2"`. Added to every data table from this wave as a `.wave` column. |
| `label` | No | A human-readable label; usually the same as `wave`. Added as `.wave_label`. |
| `wave_dir` | Yes | The folder the wave's globs are resolved against, relative to the dataset folder. Use `"."` when all waves share the dataset root and the wave is distinguished by filename; use the subfolder name when each wave has its own folder. |
| `resources` | Yes | A list of resources for this wave, structured exactly as above (minimum one). |

Repeat the whole wave block for every wave, and give resources wave-specific names (`data_2010`, `data_2014`, …) so the tidier can tell them apart.
See the [cookbook](spec-cookbook.qmd) for both wave layouts.

## Validating your spec {#validating-your-spec}

Once `dataset.yaml` is filled in, check that it reads the real files:

```r
targets::tar_make(names = raw_BPIPD_9999) # your dataset id
```

If this completes without error, the spec is reading the data.
If not:

1. Do the paths in every `glob` match the real folder structure exactly (capitalisation, spaces, subfolders)?
2. Does `reader` match the file extension (table above)?
3. Does every `role: data` resource have a `reader`?
4. Is every `role: docs` resource free of `reader`/`sheet`/`range`?
5. Do Excel resources name the correct `sheet`?
6. If the same filename exists in two folders, is the glob specific enough to match only the authoritative one?
   (When in doubt about which copy is authoritative, ask the team lead.)

The repository's test suite additionally checks every committed spec's structure (ID matches the folder, valid `status`, tidier file exists) on every pull request — but only a local `tar_make()` proves the globs match the real files, because CI has no access to the data.
