---
title: "The dataset.yaml spec"
---

Every dataset has a spec file at `harmonisation/datasets/BPIPD-<id>/dataset.yaml` that tells the pipeline what files exist, where they live, and how to read them.
This page is the field-by-field reference; for worked examples of common layouts, see the [Spec cookbook](spec-cookbook.qmd).

The shape of the file is defined by `harmonisation/dataset.schema.json`.
A minimal spec looks like this:

```yaml
dataset_id: "9999"
dataset_name: "Example Cohort Study"
status: "in_progress"

resources:
  - name: "data"
    role: "data"
    glob: "data.sav"
    reader: "spss"
```

## Top-level fields

| Field | Required? | Notes |
|-------|-----------|-------|
| `dataset_id` | Yes | The dataset number **only**, as a string of digits (e.g. `"9999"`, not `"BPIPD-9999"`). Must match the folder name. |
| `dataset_name` | Yes | The name shown next to the dataset ID on the original data folder, e.g. `"Example Cohort Study"`. |
| `status` | Yes | One of `planned`, `in_progress`, `harmonised`, `blocked`. Leave as `in_progress` until harmonisation is complete. Only `in_progress` and `harmonised` datasets are included in the pipeline. |
| `resources` | Optional | File list for datasets **without waves** (see below). |
| `waves` | Optional | File lists per wave for datasets **with waves** (see below). |
| `notes` | Optional | Free-text notes about the dataset. |

A dataset normally has *either* `resources` *or* `waves` — not both, unless there are dataset-wide files that sit outside any single wave (such as a shared codebook), in which case a top-level `resources` list can coexist with `waves`.

## Resources

Each item in `resources` describes one file — or one group of related files — in the dataset.

| Field | Required? | Notes |
|-------|-----------|-------|
| `name` | Yes | A short label, e.g. `"data"`, `"codebook"`. Names must be unique across the whole file (including across waves and across roles) — this is the name the tidier uses to access the table, and the pipeline errors on duplicates. |
| `role` | Yes | One of `data`, `codebook`, `docs` (see below). |
| `glob` | Yes | The filename, including the folder path if nested (see below). |
| `reader` | **Only for `role: data`** | How to read the file (see the reader table). |
| `sheet` | Excel only | Sheet name or 1-based index. |
| `range` | Excel only, optional | A cell range, e.g. `"A1:H200"`. |
| `table` | Access (`mdb`) only | The table to read; may be omitted when the file contains exactly one table. |
| `object` | `rda` only | The object to extract; may be omitted when the file contains exactly one object. |
| `col_names` | `csv`/`csv2`/`tsv` only, optional | Whether the first row is a header. Default `true`; set `false` for a headerless file, whose columns are then named `X1`, `X2`, … for the tidier to rename (see [Headerless delimited files](#headerless-delimited-files)). |
| `col_positions` | `fwf` only, **required** | Path to the fixed-width layout CSV, relative to the dataset's spec folder (see [Fixed-width files](#fixed-width-files)). |
| `encoding` | Optional | Character encoding of the source file, e.g. `"latin1"`. Accepted by `csv`, `csv2`, `tsv`, `fwf`, `stata`, `spss` and `sas`; any other reader refuses it. |

### Roles

| Role | Use for |
|------|---------|
| `data` | The actual dataset file(s) — `.sav`, `.dta`, `.csv`, `.xlsx`, etc. **Requires a `reader`.** Read into R and passed to the tidier. |
| `codebook` | Data dictionaries and variable documentation. The file path is recorded for reference, but the file is **not read into R**. |
| `docs` | Anything else: technical reports, user guides, questionnaires, syntax files. Paths only — must not have a `reader` or any reader option (`sheet`, `range`, `table`, `object`, `col_names`, `col_positions`, `encoding`). |

### Globs

`glob` locates files relative to the dataset's root folder (or the wave's `wave_dir`, for wave resources).

- **File in the dataset folder**: `glob: "data.sav"`
- **File nested in a subfolder** — include the full path: `glob: "Survey Files/2000/data.dta"`
- **Wildcards** — `*` and `?` are supported and can match several files: `glob: "stata/*.dta"`.
  Each match becomes its own table, named `<name>__1`, `<name>__2`, … (see [Tidy scripts](tidy-scripts.qmd#what-the-tidier-receives)).
- **Several named files in one resource** — as a YAML list:

  ```yaml
  glob:
    - "file1.sav"
    - "file2.sav"
  ```

  or with brace notation: `glob: "{file1.sav,file2.sav}"`.
  Both are equivalent; pick whichever reads better and stay consistent within a file.

Paths must match the real files exactly, including capitalisation and spacing — typos in globs are the most common reason a spec reads nothing.

## Readers

`reader` is required whenever `role: data`.
Pick it from the file *format*, which usually but not always follows from the extension: a `.dat` or `.txt` file can be comma-separated, tab-separated, headerless or fixed-width, so open it and look before choosing.

```{r}
#| label: reader-table
readers <- tibble::tribble(
  ~Reader, ~`File types`, ~`Extra keys`,
  "csv", "`.csv` (comma-separated)", "`col_names`, `encoding` (both optional)",
  "csv2", "`.csv` (semicolon-separated, comma decimal)",
  "`col_names`, `encoding` (both optional)",
  "tsv", "`.tsv` (tab-separated)", "`col_names`, `encoding` (both optional)",
  "fwf", "`.dat` / `.txt` (fixed-width, no delimiter)",
  "`col_positions` (required), `encoding` (optional)",
  "stata", "`.dta` (Stata)", "`encoding` (optional)",
  "spss", "`.sav` (SPSS)", "`encoding` (optional)",
  "por", "`.por` (SPSS Portable)", "",
  "sas", "`.sas7bdat` (SAS)", "`encoding` (optional)",
  "rds", "`.rds` (single R object)", "",
  "rda", "`.rda` / `.RData` (saved R workspace)",
  "`object` (optional if the file holds one object)",
  "parquet", "`.parquet` (Apache Parquet)", "",
  "excel", "`.xlsx` / `.xls` (Excel)",
  "`sheet` (recommended), `range` (optional)",
  "mdb", "`.mdb` / `.accdb` (Microsoft Access)",
  "`table` (optional if the file holds one table)"
)
knitr::kable(readers)
```

```{r}
#| label: reader-drift-check
#| include: false
# Fail the site build if the table above drifts from the schema's enum.
schema <- jsonlite::read_json("../harmonisation/dataset.schema.json")
schema_readers <- unlist(
  schema$definitions$resource$properties$reader$enum
)
stopifnot(setequal(readers$Reader, schema_readers))
```

For **Excel data files**, always set `sheet` — open the file in the original data folder to check the exact sheet name:

```yaml
- name: "questionnaire"
  role: "data"
  glob: "questionnaire.xlsx"
  reader: "excel"
  sheet: "Data"
```

(`reader` and `sheet` are only meaningful on `role: data` resources — codebooks are recorded by path but never read.)

::: {.callout-warning}
`.rda` files use `reader: "rda"`, not `rds`.
An `.rds` file stores exactly one object; an `.rda` file is a saved workspace that can hold several, which is why `rda` takes an optional `object` key.
:::

::: {.callout-warning}
`.por` files are SPSS *Portable* files and use `reader: "por"`, not `spss`.
`reader: "spss"` fails on them with "Invalid file, or file has unsupported features".
The `por` reader has no `encoding` option; setting one is an error.
:::

### Headerless delimited files {#headerless-delimited-files}

Some exported flat files, such as Mplus `.dat` files, have no header row.
Read with the default `col_names: true`, the first participant's row becomes the column names and the table comes back one row short, and nothing errors.
Set `col_names: false` so every row is data:

```yaml
- name: "analysis_file"
  role: "data"
  glob: "analysis_file.dat"
  reader: "tsv"
  col_names: false
```

The columns are then named `X1`, `X2`, … and the tidier renames them from the study's documentation.

### Fixed-width files {#fixed-width-files}

A fixed-width file has no delimiters: each variable occupies a fixed range of character positions on every line, and the codebook says which.
Read with a delimited reader it comes back as a single character column, and nothing errors.

Use `reader: "fwf"` and point `col_positions` at a *layout CSV* committed next to the spec, conventionally at `harmonisation/datasets/BPIPD-<id>/layouts/<name>.csv`:

```yaml
- name: "survey_2007"
  role: "data"
  glob: "survey07.dat"
  reader: "fwf"
  col_positions: "layouts/survey_2007.csv"
```

The layout CSV has one row per field:

| Column | Required? | Meaning |
|--------|-----------|---------|
| `name` | Yes | Column name; must be unique. |
| `start` | Yes | First character position, 1-based, inclusive. |
| `end` | Yes | Last character position, inclusive. |
| `type` | No | Single-letter `readr` column type (`c` character, `d` double, `i` integer, `l` logical); leave blank or omit the column to let `readr` guess. |

Any other column is ignored by the reader, so keep the codebook's format string or decimal count alongside for provenance and for the tidier.
Fields must be listed in ascending order and may not overlap; a gap between fields is fine and simply skips those characters.

Derive the layout once from the study's codebook or data dictionary and commit it, with a `README.md` next to it recording how it was produced, so the layout is reviewable and reproducible.
The pipeline validates the layout before reading and fails loudly on a bad one.
It also warns when the layout runs past the end of the first record, because `read_fwf()` would otherwise return `NA` for the overhanging field without complaint.
Layout files are tracked like data files: editing one rebuilds the read.

Implied decimals (a codebook format such as `N8.3`) are **not** applied by the reader; carry the format in an extra layout column and divide in the tidier.
There is a worked example in the [cookbook](spec-cookbook.qmd#fixed-width-file).

## Waves

If a dataset has files for multiple waves (years, sweeps), use `waves` instead of a top-level `resources` list:

```yaml
waves:
  - wave: "2010"
    label: "2010"
    wave_dir: "."
    resources:
      - name: "data_2010"
        role: "data"
        glob: "survey_2010.csv"
        reader: "csv"
```

| Field | Required? | Notes |
|-------|-----------|-------|
| `wave` | Yes | The wave identifier, e.g. `"2001"`, `"sweep_2"`. Added to every data table from this wave as a `.wave` column. |
| `label` | No | A human-readable label; usually the same as `wave`. Added as `.wave_label`. |
| `wave_dir` | Yes | The folder the wave's globs are resolved against, relative to the dataset folder. Use `"."` when all waves share the dataset root and the wave is distinguished by filename; use the subfolder name when each wave has its own folder. |
| `resources` | Yes | A list of resources for this wave, structured exactly as above (minimum one). |

Repeat the whole wave block for every wave, and give resources wave-specific names (`data_2010`, `data_2014`, …) so the tidier can tell them apart.
See the [cookbook](spec-cookbook.qmd) for both wave layouts.

::: {.callout-warning}
## Wave resource names must differ

Copying a wave block and changing only `wave`, `wave_dir` and `glob` leaves every wave declaring a resource called `data`.
Resource names are the keys of `raw_dataset$data`, so `raw_dataset$data$data` would then return wave 1 and nothing else, and every later wave would be unreachable by name.

The pipeline refuses to read a spec in that state, listing each repeated name and the waves it appears in:

```
Error in `check_resource_names()`:
! Duplicate resource names in `harmonisation/datasets/BPIPD-9999/dataset.yaml`.
✖ `data` is declared 3 times (wave 2010, wave 2014, wave 2018).
ℹ Names must be unique across the whole spec, including across waves and across roles.
```

The fix is to work the wave into every name — including `codebook` and `docs` resources, which share the same namespace.
:::

## Validating your spec {#validating-your-spec}

Once `dataset.yaml` is filled in, check that it reads the real files:

```r
targets::tar_make(names = raw_BPIPD_9999) # your dataset id
```

If this completes without error, the spec is reading the data.
If not:

1. Do the paths in every `glob` match the real folder structure exactly (capitalisation, spaces, subfolders)?
2. Does `reader` match the file *format* (table above)?
   A `.dat` or `.txt` file may be delimited, headerless or fixed-width.
3. Does every `role: data` resource have a `reader`?
4. Is every `role: docs` resource free of `reader` and reader options?
5. Do Excel resources name the correct `sheet`?
6. If the same filename exists in two folders, is the glob specific enough to match only the authoritative one?
   (When in doubt about which copy is authoritative, ask the team lead.)
7. Is every resource `name` unique across the whole file, including across waves?
8. Did every value parse?
   A cell that cannot be read as its column's type would otherwise become `NA` without trace, so the pipeline refuses the file and names the column, the number of failures and the first offending value (for example `n/a` or `.` in a numeric column).
   Fix the column `type` in a fixed-width layout, or leave the column as text and recode the value in the tidier.
9. Does each table have the shape the codebook promises?
   A wrong reader can return a structurally valid table — a single column, or one row short — without erroring, so check `dim(targets::tar_read(raw_BPIPD_9999)$data$<name>)` against the documented row and variable counts.

The repository's test suite additionally checks every committed spec's structure (ID matches the folder, valid `status`, unique resource names, tidier file exists) on every pull request — but only a local `tar_make()` proves the globs match the real files, because CI has no access to the data.
