---
title: "Harmonising variables"
---

Harmonisation maps a dataset's [tidied table](tidy-scripts.qmd) onto the project [dataschema](../reference/dataschema.qmd), so that every study contributes the same variables with the same meanings.
It is entirely table-driven: you edit CSV files, not R scripts.

The files involved:

| File | Purpose |
|------|---------|
| `harmonisation/dataschema.csv` | The shared schema — one row per harmonised variable (shown in the [Dataschema reference](../reference/dataschema.qmd)) |
| `harmonisation/datasets/BPIPD-<id>/variables.csv` | How *this* dataset derives each schema variable |
| `harmonisation/lookups/*.csv` | Shared recode tables reused across studies (e.g. ethnicity) |
| `harmonisation/datasets/BPIPD-<id>/lookups/*.csv` | Large one-off recode tables specific to this dataset |

## variables.csv

One row per dataschema variable, in dataschema order.
The template copy starts with every variable set to `in_progress`; harmonising a dataset means working through the rows.

| Column | Meaning |
|--------|---------|
| `target_variable` | The dataschema variable being derived. Every non-system schema variable must have a row (`dataset_id`, `dataset_name` and `st_measure_id` are injected by the pipeline). |
| `status` | How well the study supports this variable (table below). |
| `source_columns` | Which tidied columns the mapping uses — documentation for reviewers, not executed. |
| `expression` | R code that computes the value (required unless the status says the variable can't be derived). |
| `notes` | Anything a reviewer needs: assumptions, DOIs, caveats. |
| `lookup_table` | Basename of a lookup CSV this expression uses, if any (e.g. `ethnicity` for `harmonisation/lookups/ethnicity.csv`). |
| `measure` | Optional; blank for the dataset's primary measure. A tag such as `diary` marks rows that map the screen-time variables again for an additional measure — see [Several screen-time measures](#several-screen-time-measures). |

A few illustrative rows (the column names are fictional, but the shapes mirror real mappings in the project):

```csv
target_variable,status,source_columns,expression,notes,lookup_table
participant_id,compatible,serial,as.character(serial),Assume serial is the participant ID,
wave,unavailable,,,Cross-sectional study,
data_year,inferred,,2016,Not in the data; taken from the study protocol,
sex,compatible,sex_raw,"recode_values(haven::as_factor(sex_raw), ""F"" ~ ""Female"", ""M"" ~ ""Male"")",Convert labels to harmonised values,
```

## Statuses

Statuses follow the [Maelstrom Research](https://www.maelstrom-research.org/) harmonisation vocabulary:

```{r}
#| label: status-table
statuses <- tibble::tribble(
  ~Status, ~Meaning, ~Evaluated,
  "compatible",
  paste(
    "The study variable can generate the target with an explicit,",
    "scientifically defensible transformation and without meaningful",
    "loss of meaning."
  ), "Yes",
  "partial",
  paste(
    "The study only supports part of the target construct, or the",
    "harmonised value requires a simplification that should be",
    "interpreted with caution."
  ), "Yes",
  "proximate",
  paste(
    "A categorical subtype of partial: the study captures a close but",
    "not fully equivalent category structure."
  ), "Yes",
  "inferred",
  paste(
    "The value is derived from study metadata or surrounding context",
    "(e.g. a fixed collection year from the study's publication)",
    "rather than a like-for-like study variable."
  ), "Yes",
  "incompatible",
  paste(
    "Related information exists, but it cannot be transformed into a",
    "defensible version of the target variable."
  ), "No — output is NA",
  "unavailable",
  "The study does not collect the required information.",
  "No — output is NA",
  "in_progress",
  "The mapping has not been completed yet (the template default).",
  "No — output is NA"
)
knitr::kable(statuses)
```

```{r}
#| label: status-drift-check
#| include: false
# Fail the site build if this table drifts from the pipeline's status set.
source("../R/config.R")
stopifnot(setequal(statuses$Status, bp_harmonisation_status_values()))
```

Rows with an "evaluated" status must have an `expression`; rows with the other three statuses are skipped and the variable is filled with typed `NA` values for that dataset.
How to choose between the statuses, including incompatible versus unavailable, is set out under [Harmonisation conventions](#harmonisation-conventions).

## Writing expressions

An expression is a single R expression evaluated against the tidied table.
Its result is recycled to the table's row count (so a constant like `2016` is fine) and cast to the schema's declared type.
Inside an expression you can reference:

- **columns of the tidied table** directly by name (`serial`, `sex_raw`);
- `tbl` or `analysis_base` — the whole tidied table, if you need it;
- `spec` — the parsed `dataset.yaml` (e.g. `spec$dataset_id`);
- `lookup_<name>` — any lookup table named in the row's `lookup_table` column (e.g. `lookup_ethnicity`).

Expressions run in a **sandboxed environment**, not your R session.
Bare function calls resolve from base R plus this curated allowlist:

```{r}
#| label: allowlist-table
source("../R/helpers.R")
source("../R/harmonise/harmonisation_functions.R")
fns <- bp_harmonisation_functions()
origin <- vapply(
  fns,
  function(fn) {
    env_name <- environmentName(topenv(environment(fn)))
    if (env_name %in% c("R_GlobalEnv", "")) "pipeline helper" else env_name
  },
  character(1)
)
allowlist <- tibble::tibble(
  Function = paste0("`", names(fns), "()`"),
  From = origin
)
knitr::kable(allowlist)
```

Namespaced calls such as `haven::as_factor()` also work for anything not on the list.
If an expression calls a bare function that is neither base R nor allowlisted, the test suite fails with a message telling you to either namespace it or add it to `bp_harmonisation_functions()` (`R/harmonise/harmonisation_functions.R`) — extending the allowlist is a one-line change, reviewed like any other.

Typical recodes use `recode_values()` with formula syntax:

```r
recode_values(
  haven::as_factor(sex_raw),
  "F" ~ "Female",
  "M" ~ "Male"
)
```

## Lookup tables

Three tiers, from smallest to largest:

1. **Inline** — short, study-specific recodes stay in the expression itself (like the `sex` example above).
2. **Shared lookups** (`harmonisation/lookups/*.csv`) — recurring recodes that should be reviewed centrally, such as ethnicity or maternal education.
   These must have the columns `dataset_id`, `target_variable`, `source_value`, `harmonised_value`, `mapping_status`, `notes`, with one row per source value per dataset.
3. **Dataset lookups** (`harmonisation/datasets/BPIPD-<id>/lookups/*.csv`) — mappings too large to read inline but not shared across studies.
   At minimum they need `source_value` and `harmonised_value` columns.

To use one, put its basename in the row's `lookup_table` column and call `lookup_values()` with the matching `lookup_<name>` object.
For example, an `ethnicity` row with `lookup_table` = `ethnicity`:

```r
lookup_values(
  haven::as_factor(ethnicity_raw),
  lookup_ethnicity,
  dataset_id = spec$dataset_id,
  target_variable = "ethnicity"
)
```

The `dataset_id`/`target_variable` arguments filter a shared lookup down to this dataset's rows; omit them for a dataset-specific lookup.
A lookup name must exist in exactly one of the two locations — the validator errors if it is missing from both or present in both.

## Several screen-time measures

A study sometimes measures screen use more than once per occasion: a questionnaire and a time-use diary, or a parent report and a child report.
The harmonised table keeps one row per participant per wave *per measure*, and the system variable `st_measure_id` says which measure a row carries.

The untagged rows of `variables.csv` describe the dataset's primary measure, and every row they produce has `st_measure_id = "primary"`.
An additional measure is a block of rows tagged in the `measure` column with a short lower-case token (`diary`, `parent`, `child`, `device`):

```csv
target_variable,status,source_columns,expression,notes,lookup_table,measure
st_measure_type,inferred,,"""Continuous""",Diary slots are 10 minutes,,diary
st_measure_name,inferred,,"""Time use diary""",,,diary
st_responder,inferred,,"""Self""",The child completes the diary,,diary
st_watch_wd,compatible,tud_tv_mins_wd,tud_tv_mins_wd / 60,Minutes to hours,,diary
```

For each tag the harmoniser evaluates the tagged rows against the same tidied table and emits one extra row per participant-wave, with `st_measure_id` set to the tag.
Only the screen-time variables can be tagged; every other variable in those rows is copied from the primary rows.
Screen-time variables the measure does not map are `NA` in its rows rather than inherited, and a participant-wave with no screen-time quantity for the measure gets no row at all.
Every measure block must map `st_measure_type`, `st_measure_name` and `st_responder`, so the rows say what instrument and respondent they came from.
On primary rows, the harmoniser sets `st_responder` to `NA` wherever every screen-time quantity is missing, because no one reported screen use there; `st_measure_type` and `st_measure_name` describe the instrument and stay filled.

The tidier stays one row per participant-wave: the extra measure arrives as extra columns (`tud_tv_mins_wd`, `tv_hours_parent`), and the mapping decides which measure they belong to.
Analyses that want the old one-row table filter on `st_measure_id == "primary"`; analyses that compare instruments use `st_measure_name` and `st_responder` as covariates.

## Harmonisation conventions

These rules apply across rows; rules for a single variable sit in its dataschema notes.
Dataschema notes refer to the named rules below in bold (for example the **SDQ rule**).

### Statuses

- Choose the status in this order: unavailable if nothing on the row's construct is in the supplied data; incompatible if an item on the construct is in hand but cannot become the row; inferred if the value is a constant or descriptor taken from documentation, file or wave labels, or row position; otherwise compatible, partial or proximate by fidelity.
- Incompatible covers items the schema routes to another row, items already used on another row and items asked only of a subsample; unavailable is only for 'nothing related in hand' or 'Not in the supplied extract.'
- Typical reasons for incompatible are a wrong metric (frequency, yes/no, rank, most-used choice or ownership instead of duration), a wrong rating format (truth, agreement, intensity or distress instead of frequency), a wrong respondent and bundled content.
- An incompatible row names the rejected item in `source_columns`, and its note says why it fails and which row holds it.
- A row that only adds a weekday/weekend, device or at-school split the study never made is unavailable, whatever the parent row's status ('No weekday/weekend split.', 'No device split; see <row>.', 'No item on use at school.').
- A named-instrument row is unavailable when that instrument, or a version of it, was not given to that informant ('<instrument> not administered; see <row>.'), and incompatible only when a version was given but cannot be scored.
- A generic scale row with only single items, or only an item subset of a broader instrument not given for the construct, is incompatible, and its note points to any row that holds them.
- A composite that can be derived at analysis from mapped rows is unavailable ('Derive at analysis from <rows>.').
- A named instrument given only to another informant leaves this informant's row unavailable.
- A single-item `_self` row with only a parent- or teacher-rated analogue is incompatible, item named.
- Compatible needs a note that describes no loss; a note saying 'only', 'undercount', 'narrower', 'bundled' or 'may include' means partial at best.
- A schema note that sends an item to another row forbids mapping it here.
- Partial covers part of the construct, pooled or coarser sources, subsets on named rows, and lifetime or unstated recall.
- Proximate is only for categorical rows whose sole loss is the category structure; binary and double rows use partial, and so does a row that loses both scope and categories.
- Inferred is for constants and for descriptor rows (`st_measure_type`, `st_measure_name`, `st_responder`, every `*_measure` row), unless a source column records the value directly, such as a who-answered variable (compatible).
- A descriptor row whose score row is unavailable is unavailable too ('See <row>.').
- A parent's or carer's own characteristic used for the child is proximate.
- A value derived from a recorded variable is never inferred.
- Coverage never sets the status: a row mapped for some waves, ages or grades takes the status its mapped values deserve, and its note starts with the coverage ('2018-2023 only.').
- A dataset that mixes waves with per-row values and waves with assigned values takes the weaker status and names the affected waves.
- A bounded recall window (two weeks to 12 months, or 'this school year') leaves the status unchanged, even when it changes between waves; a lifetime or unstated window is partial.
- On the binary prevalence rows for suicidal ideation, suicide attempt and depression screening, any recall other than 12 months is partial.
- A weaker item fills waves the better item missed only on score rows standardised within wave (partial, waves in notes).
- Yes/no prevalence rows keep the better series alone; they never splice in a different recall window or a pooled item.
- Every unavailable or incompatible row has a one-line note.
- Standard notes are 'Not collected.', 'Not in the supplied extract.', 'No weekday/weekend split.', 'No device split; see <row>.', 'No item on use at school.', '<instrument> not administered; see <row>.', 'Derive at analysis from <rows>.', 'Not an Australian dataset.', 'Hukou not asked.', 'Not a randomised trial.', 'Single collection occasion.', 'Not a survey sample.' (survey design rows) and 'No cognitive testing.' (cognitive rows in studies that gave no test)
- An unavailable note never describes an item in hand and never lists the study's other instruments.
- `source_columns` lists tidied column names, comma-separated, with no ranges and no raw-file names.
- Notes quote the item wording and record the reporter, the waves and the recall period.
- A row accepts any informant unless its name or description says otherwise; record the informant.
- Notes state the rule applied and never cite another dataset as precedent.
- A choice settled for one dataset only stays in that dataset's notes.
- Dataschema text is public: it never names real datasets or their source columns.
- Before finishing, re-read every note against its status.

### Screen time

- Every `st_*` quantity is hours per day as a double.
- Minutes / 60; hours and minutes h + m / 60; weekly totals / 7; a total for both weekend days / 2.
- Clear negative and sentinel codes to NA before any arithmetic.
- Bands: 'none' = 0, closed bands at their midpoint, 'less than 1 hour' = 0.5, sub-hour bands at their midpoint ('less than 30 minutes' = 0.25), point labels ('about 2 hours') at the stated value.
- An open top band takes the lower bound its label states ('5 or more' = 5, 'more than 7 hours' = 7).
- A point label below the top band ('about 4 hours') is read as a band of plus or minus half an hour, so 'more than 4 hours' takes 4.5.
- Write the top-band value into the row's notes.
- Recode raw bands per wave where band sets differ, and use study-derived hours only if they follow these rules.
- Weekday and school-day items map to `_wd`; weekend and non-school-day items, holidays included, map to `_we`.
- The unsuffixed row is (5 × wd + 2 × we) / 7, and only when both are mapped for the same wave.
- Weekday-only items fill `_wd` only; the unsuffixed row is incompatible ('Weekday items only; no usual-day value.') and `_we` is unavailable.
- Usual-day items fill the unsuffixed row only; `_wd` and `_we` are unavailable ('No weekday/weekend split.').
- With no day split, `_wd` and `_we` are unavailable even when the unsuffixed row is incompatible.
- `*_school` rows take use while at school only; an item about a whole school day goes to `_wd`.
- **Non-users:** 0 when a filter establishes non-use (routed past the item, 'not at all', or no device with use unanswered); NA when not asked or not answered; a filter never overrides a reported duration.
- Keep values above 24 hours; they are excluded at analysis, with a sensitivity analysis that keeps them.
- Minutes on a day of use are multiplied by the share of days used (days per week / 7) and marked partial; minutes with no frequency answer are NA.
- Frequency-only, rank, most-used, yes/no and ownership items give no hours, so the row is incompatible, item named.
- Sum only sub-categories of one row (DR-0002): devices within gaming, sources within watching, gaming and non-gaming within computer, activities within one device, locations within one content row.
- Write 'Sum of A, B (DR-0002)' in the notes of a summed row.
- Never sum across watching, gaming, computer, phone, social media or internet, including into `st_total`, whoever did the summing.
- Screen-time sums use `+`, with routed-out parts written as explicit zeros; `sum_nonmissing()` only where the instrument scores blanks as zero.
- A skipped part is never scored as zero, so it leaves the sum NA.
- An item may feed its umbrella row as a DR-0002 part, but never stands in for a broader or sibling row: texting is not `st_phone`, phone gaming is not `st_phone`, internet time is not `st_total`.
- A single-activity item mapped to its content row leaves the device row incompatible.
- A bundled item maps to one row as partial, and every other row it touches is incompatible, naming the item and the row it went to.
- Social media with messaging goes to `st_socmedia`.
- Gaming with other non-school computer use goes to `st_game`, and `st_comp` is incompatible.
- Gaming or computer items whose examples name social media never map to `st_game*`.
- Phone with tablet goes to `st_phone` or `st_game_phone` (partial), and the tablet row is incompatible.
- Computer with tablet as one response option goes to `st_tablet` or `st_game_tablet` (partial), and the computer row is incompatible.
- Residual 'other online or device use' goes to `st_internet`.
- A bundled or residual item is never added to a row that already has a direct item, nor spliced onto a row that a different item fills in other waves; name it as unused.
- A social-media item whose examples name video-sharing platforms is not bundled with watching.
- Watching pooled with gaming, computer or internet use goes to `st_total`.
- Under a pooled multi-device gaming item, the device rows are unavailable ('No device split; see st_game.'); pooling activities within one device is bundling and stays incompatible.
- `st_game` covers gaming on any device.
- A gaming item that names devices and leaves out phones and tablets is partial whatever its year; an item naming no device is compatible.
- An item limited to the TV set (broadcast TV, or videos and DVDs on the set) is partial on `st_watch`; the test is the device, not the word 'videos'.
- A computer item that excludes games goes to `st_comp_nongaming`; other purpose-restricted device items are partial.
- Internet time on one device maps to that device's row as partial.
- Study or homework time that does not name screens is incompatible, except during fully remote schooling (partial).
- A row that splices instruments with a documented level shift is partial.
- Screen-time doubles never use proximate.
- `st_measure_type` records the format behind the row's core `st_` variables.
- `st_measure_name` takes one controlled label: a published instrument name, 'Time use diary', 'Custom' or 'Unknown'.
- `st_measure_type` and `st_measure_name` are filled wherever the instrument was administered; the harmoniser sets only `st_responder` to NA on primary rows with no `st_` value.
- One responder per row, including any filter that sets a value to zero; responders are never combined within a row or a variable.
- The study's main series is primary, and a second responder, instrument or diary on the same occasion is a tagged block.
- Block tags are `diary`, `parent`, `child`, `device` and `recall`, and every block maps `st_measure_type`, `st_measure_name` and `st_responder`.
- A `recall` block covers the same occasion with a different reference period; its other variables belong to the occasion, so concurrent analyses exclude it.
- Waves, and country-cycles of optional modules, with no disaggregated screen item are dropped in the tidier, so expressions need no wave conditions.

### Scoring scales

- Higher means more of the construct the description names; reverse source scales to match.
- Rebase in the expression (subtract the item count for 1-based items; reverse as (min + max) - x) and record both in notes.
- Set sentinel and off-scale item codes to NA (`na_if()`, `%in%` valid codes) before scoring, and check the observed range against the published range.
- Before mapping a study-supplied total, check its keying (every item correlates positively with the total) and rescore from items if it fails.
- A named-instrument row takes the full published form on its published scoring, or a study composite equated to that metric (partial).
- The same items on a non-standard response scale are partial on the named row.
- Short forms, item subsets and adapted forms go to the generic row, and the named row is incompatible.
- Exceptions: `mh_dep_scl_self` takes SCL-90-R short forms, the RCADS-25 depression items map to `mh_dep_rcads_self`, and mean-item CBCL rows take item subsets as partial.
- A generic ('any instrument') row takes a multi-item scale administered for the construct.
- On a generic row the status reflects construct fit only; short forms, subsets and adapted scoring are compatible when the `*_measure` row describes them.
- When a named self-report row is mapped, repeat its expression on the generic row.
- Standardise generic rows, single-rating rows and raw-score rows within dataset at analysis, and within wave where the scale or form changes.
- `*_measure` rows read '<instrument>, <k> of <K> items, <a>-<b> item scoring (range <min>-<max>), <sum|mean>', then the reversed items and any change of response format or recall period.
- `*_measure` rows are inferred, entity study, filled in waves (or country-cycles) where the scale was given, a constant when every wave fielded it, and NA elsewhere, never 'Unknown' as a filler.
- A `*_measure` row says 'Unknown' only when a score exists but its instrument cannot be identified.
- Use a study-supplied score as released and record its missing-item rule ('rule not documented' if unknown).
- Sums computed in the expression are complete-case (`+`), unless the manual prescribes proration.
- **SDQ rule:** items scored 0-2 (not true 0, somewhat true 1, certainly true 2), items 7, 11, 14, 21 and 25 reversed, subscale = mean of answered items × 5 rounded half up with `floor(x + 0.5)`, NA when fewer than 3 of 5 are answered.
- Where the SDQ items are released, check the study's subscale against them: keep it if it follows the SDQ rule, otherwise gate or recompute it from the items.
- Where the SDQ items are not released, use the study's subscale after checking its range and reversals, and say in notes that the missing-item rule could not be verified.
- An SDQ subscale is incompatible when fewer than 3 of its items were administered, and total difficulties is scored only when all four difficulty subscales are.
- Pool the SDQ 2-4 and 4-17 parent forms and record the form version.
- SDQ self rows keep every age administered and record the age range and under-11 share.
- SDQ externalising and internalising are unavailable ('Derive at analysis from the subscale rows.') unless only the composite is released.
- **CBCL rule:** mean of the answered 0-2 items of the form's scale, NA when fewer than 80% of administered items are answered, codes outside 0-2 set to NA.
- CBCL composites take the mean over the union of component items, gate each component at 80%, and follow the form's published grouping.
- `beh_cbcl_measure` records the form, items administered of the published total per scale, the scoring and the missing rule, one value per wave where forms differ.
- PedsQL items are transformed 0-4 to 100-0 and averaged when at least half are answered.
- BPAQ-SF rows take the published items, never a study's own syntax, and need all three items.
- Subject composites (grades, teacher ratings) average every academic subject reported, not arts, PE or ICT, and need every component (partial).
- Respondent suffixes are literal, respondents are never coalesced, and the CBCL and BASC-2 rows are parent report only.
- Cognitive rows take the age-normed score the study releases: standard score (mean 100, SD 15), T-score (50, 10) or scaled score (10, 3).
- Raw, ability or number-correct scores map only when no normed form exists, as partial; percentiles, age-equivalents and normative classifications never map.
- Never mix score forms within a row across waves.
- NIH Toolbox rows take the age-corrected standard score (uncorrected only as partial), and the BASC-2 executive-function row takes the T-score; never the fully corrected T-score.
- BAS II subtests are given individually, so there is no GCA for `cog_iq_fullscale`; map T-scores, with ability scores only as a partial fallback.
- NEPSY rows take the same-named subtest from either edition; record the edition and age at testing.
- Only the child's own test maps; tests the caregiver took never map to child rows.
- Parent-rated executive-function scales run higher = more problems and never pool with performance tasks.
- Plausible-value scales use the first plausible value only.
- Achievement tests take one domain per row (`acc_academachment_stdtest` records its domain); age-normed standard or scale scores are compatible and raw scores partial.
- Record the test, subtest and score type, and standardise test scores within dataset and wave at analysis.
- Social-support rows take only items on support the child receives or can call on (help, understanding, someone to talk to or confide in).
- Parenting behaviours and family involvement are incompatible on the support rows, and a published support subscale that mixes in companionship items is partial.
- School belonging takes only scales presented as belonging or connectedness; teacher- or classmate-support scales and school-environment quality-of-life subscales are incompatible.

### Single items

- **Frequency-item rules** (`mh_*_self` rows with Always to Never): self-reported frequency ratings only; truth, agreement, intensity and distress ratings are incompatible, item named.
- Map anchors by wording: never, none, not at all -> Never; almost never, hardly ever, seldom, less than 1 day -> Rarely; sometimes, occasionally, some of the time, several days -> Sometimes; often, more than half the days -> Often; always, almost always, all of the time, nearly every day -> Always.
- An anchor steps down one label when the source has an intensified anchor above it ('quite often' under 'very often' -> Sometimes; 'most of the time' under 'always' -> Often), and a top anchor above a source 'often' -> Always.
- Day counts without wording map by share of a known recall window: 0 -> Never, up to 10% -> Rarely, up to 35% -> Sometimes, up to 70% -> Often, above 70% -> Always.
- Labels the source cannot reach stay unused and are named in notes; a source with fewer than five levels is proximate; a wording mismatch is partial.
- Record the recall period on every frequency row; it does not change the status there.
- An item also scored in a scale may fill a single-item row when it is self-reported, frequency- or presence-rated and matches in wording; note 'also scored in <row>'.
- Score rows take only published scale or subscale scorings, never an ad hoc sum of another instrument's items.
- **GAD-7 crosswalk:** not at all, several days, more than half the days, nearly every day -> Never, Sometimes, Often, Always (PHQ items the same).
- CES-D day bands <1, 1-2, 3-4, 5-7 -> Rarely, Sometimes, Often, Always; all, some, none of the time -> Always, Sometimes, Never.
- A measure of another person (a parent's own stress or diagnosis) is unavailable on a `_self` row.
- Always/Usually/Sometimes/Never rows take frequency scales only; degree, importance and agreement scales are incompatible.
- On those rows, map anchors by wording with the same positional step; a four-level source whose labels do not match maps by rank, proximate, with the anchors in notes.
- 'All/Most/Some of the time' = Always/Usually/Sometimes, and a negatively worded item is never reversed onto these rows.
- **Five-level collapse** onto Very good, Good, Average, Below average: top -> Very good, second -> Good, middle -> Average, two lowest -> Below average (proximate), from the full-resolution source item.
- The four HBSC-SCL single-item rows are named-instrument rows: unavailable ('HBSC-SCL not administered; see <row>.') wherever neither the checklist nor the adaptation in the row's note was given, even when a related item is in hand.
- On school pressure, worry about school performance and test anxiety are related items (incompatible, item named), and a five-point pressure scale maps its top two levels to 'A lot'.
- A behaviour item that places a device in the bedroom is partial on `fam_bedroomscreen`, and a parent's technical time limit is partial on `fam_screenrules`.
- Binary behaviour rows are TRUE for any occurrence (the lowest response above never) and FALSE for never; yes/no items map directly.
- A multi-item composite is TRUE if any answered item is positive, FALSE if every answered item is never, and NA if none is answered; study missing codes count as unanswered.
- Use self-report where it exists; a parent-reported occurrence item is partial.
- An item limited to one sub-type or setting of the behaviour is partial; school-based items are compatible on the bullying rows.
- Truth- or agreement-rated items (SDQ, CBCL, 'describes this child') never map to binary or frequency rows; frequency-rated instrument items (BASC-2, PedsQL) are judged on content.
- School rows (truancy, school and schoolwork satisfaction) are NA for respondents not enrolled in school or enrolled beyond secondary school.
- Single-rating rows (life, domain and happiness ratings) record the scale points and anchor wording and reverse scales where the lowest number is best.
- Keep one response scale per dataset per single-rating row where possible; where it changes between waves, standardise within dataset and wave.
- Faces scales (completely happy to not at all happy) count as satisfaction ratings, and life as a whole always goes to `wb_lifesat_singleitem_self`.
- Diagnosis rows: Never = never told, Previously = told but not current, Currently = told and current; treatment or medication as the proxy for current is partial; items not attributed to a health professional are incompatible.

### Categorical recodes

- Emit `allowed_values` exactly, case included.
- Demographic and SES rows emit the literal "Unknown", never NA, on every row of a mapped dataset, including waves where the item was not asked and nothing is carried (`default = "Unknown"` or `coalesce(..., "Unknown")`).
- `state` is the exception (NA outside Australia), and so is `ses_hukou` (NA outside mainland China or where not asked).
- Academic, behaviour and mental-health categorical rows use NA, never Unknown; don't know, not graded, pass/fail and not applicable codes are NA.
- A source code no document defines maps to Unknown on demographic rows and NA elsewhere, never to a real category.
- A category merging a real value with Other or unknown maps to the real value (Female or unknown -> Female; Other or unknown -> Other), proximate, merge named; a merged Other/prefer-not-to-say maps to Unknown.
- A group with no schema category folds into Other and is proximate.
- A Hispanic-origin item sets Hispanic only in US and Canadian samples; elsewhere the race, colour or ethnic-group item decides.
- Minority, indigenous and immigrant background are judged against the country of data collection; multi-country data use a dataset lookup keyed on site and source value, with the reference group in notes.
- Derived rows use the same item, lookup, free-text resolution and carry-forward as their source: minority and indigenous follow ethnicity.
- Released 'other, specify' text is resolved through a dataset lookup; a bare nationality is Unknown.
- A participant-level characteristic is never a constant from study context; country is the exception, and so is `born_in_country` Yes (inferred) where recruitment at birth in the country guarantees it.
- Birthplace rows use the UK as a whole and count Hong Kong and Macao in their own right.
- Recurring vocabularies use one shared lookup each (ethnicity, education), and all three parent-education rows call the same lookup.

### Age and time

- Completed years take +0.5 (13 -> 13.5); a completed-year band a-b takes (a + b + 1) / 2; an open band takes its inclusive bound + 0.5.
- The +0.5 rule is for age only; open screen-time bands take their floor.
- Survey year minus birth year is adjusted by the fieldwork midpoint, and decimal age is derived where birth and interview months are released.
- A single-age sample with no age item takes the centre of the sampled cohort's age at fieldwork, inferred.
- Single completed years and survey-year differences are compatible; multi-year bands, open bands holding more than about 5% of rows and design-assigned ages are partial or inferred; a birth date against an assumed reference date is partial.
- Ages above `max_value` are kept, eligibility is applied at analysis, and tidiers drop rows only to define the study population and the waves each participant attended.
- `data_year` takes the per-row interview year where released (compatible) or the survey's own year label (compatible); a year recoded from wave labels or documentation is inferred.
- A missing per-row date takes the wave's modal fieldwork year; a few such fills leave the row compatible, with the count in notes, and inferred is for waves whose year is assigned wholesale.
- Constants cite fieldwork or design documentation with dates, never a publication year.
- A single-occasion dataset leaves `wave` NA (unavailable); repeated cross-sections use the four-digit survey year.

### Multiple waves and carry-forward

- **Carry-forward:** sex, ethnicity, indigenous status, immigrant background, `born_in_country` and derived minority are carried within participant through one allowlisted helper in the expression; observed values stay, and missing or not-asked waves take the earliest observed value in wave order.
- Sex items take the participant's modal observed value at every wave, ties Unknown; gender-identity items stay as recorded.
- Parent education is always carried; household SES and family structure are carried across gaps of 12 months or less (inclusive) and are otherwise Unknown; this fill is done in the tidier.
- A follow-up wave with no age item takes the nearest recorded age plus the documented interval, filled in the tidier, partial.
- Waves and country-cycles with no disaggregated screen item are dropped in the tidier.
- Tidiers drop a participant's rows for waves the participant did not attend.
- Survey design values fixed at sampling (stratum, cluster, design weight) are carried within participant to every wave without a wave prefix, so a participant stays in one stratum and cluster; a carried design weight is partial.
- The weight is the study's core release weight, and schools in a randomised trial are not survey clusters.
- Wave coverage alone never lowers a status; list the waves in notes.
- A second report on the same occasion is a tagged measure block, never a wave.
- A `*_measure` string appears only in waves where its scale was given.

### Sleep

- Time of falling or going to sleep to waking counts as hours asleep (compatible unless banded); bedtime or lights-out to waking is partial.
- Never subtract a separate sleep-latency item.
- An open sleep band takes the bound its label states, or half an hour beyond a neighbouring point label ('less than 6 hours' next to '6 hours' = 5.5).
- The share of answers in an open band does not lower the status.
- Keep durations under 4 or over 16 hours; they are screened at analysis.

### SES

- **Household SES cuts:** Low, Medium and High split the unweighted sample 40/20/40 within wave, and within country or site where the measure or its bands are country- or site-specific; at or below the 40th percentile is Low and above the 60th is High.
- A banded or discrete source classifies each band by its ridit in the same frame (below 0.4 Low, 0.4 to 0.6 Medium, above 0.6 High); if no band falls in 0.4 to 0.6, the band nearest 0.5 is Medium.
- A fixed class scheme keeps its standard three-way collapse (proximate), and subjective standing is cut at its labelled anchors with the midpoint as Medium.
- Income, income-to-needs, affluence or a composite index is compatible when every level lies within 10 points of 40/20/40 in every wave, and proximate otherwise.
- Multi-country repeated cross-sections cut within country and cycle.
- Area SES uses the index's own national quintiles, never sample quantiles.
- School-level indicators (school type, fees, funding, track) map to neither SES row; mark them incompatible with a one-line note.
- `ses_household_measure` reads '<type>: <measure>', with the type from a fixed list.
- Parent education uses one vocabulary with ISCED anchors; a band spanning two levels goes to the level holding most of its members (proximate), and tertiary study begun but not completed is Post-secondary.
- Parents are whichever mother and father figures the source records, with the row partial when one is systematically unobserved.
- Categorical SES rows are compatible when every level matches the schema's meaning and proximate for coarser, lumped or stand-in sources (school location, subjective standing, occupational class); they never use partial, and constants are inferred.
- Urbanicity follows a national typology on its own thresholds; raw size bands use under 5,000 Rural, 5,000 to under 100,000 Suburban and 100,000 or more Urban.
- An urban constant is allowed only when the design guarantees urban residence at the measured wave.

## Keeping variables.csv in sync with the dataschema

When the dataschema gains a variable, every dataset's `variables.csv` needs a matching row.
Don't do this by hand:

```r
targets::tar_source()
sync_harmonisation_vars() # all datasets, plus the new-dataset template
```

This appends any missing variables as `in_progress` and reorders rows to dataschema order, keeping any measure-tagged rows after the primary rows.
It covers the template under `harmonisation/templates/dataset/` as well, so a newly copied dataset spec starts with the full set of variables.
The validator (and therefore CI) rejects a `variables.csv` with missing, duplicate, or unknown target variables.

## Checking your expressions

Before running the pipeline or the test suite, you can check every `expression` in one go:

```r
targets::tar_source()
check_harmonisation_exprs()
```

It returns one row per problem — an empty tibble means everything is fine — with the columns `variables_file`, `target_variable`, `measure`, `status`, `issue`, `message`, `suggestion`, and `fixed`.
There are four kinds of `issue`:

| Issue | Meaning |
|-------|---------|
| `parse_error` | The expression is not valid R and will fail the moment it is evaluated. |
| `unknown_function` | It calls a function that is neither base R, allowlisted, nor namespaced (see [Writing expressions](#writing-expressions)). |
| `missing_expression` | The status says the row is evaluated, but no expression was given. |
| `unformatted` | It is valid, but [Air](https://posit-dev.github.io/air) would format it differently; `suggestion` holds the reformatted text. |

The first three are what CI enforces, so a clean run here means the expressions will survive the test suite.
Formatting is advisory only and is never enforced by CI — pass `check_format = FALSE` to skip it, which also avoids needing Air installed.
Air ships inside a VS Code extension in this devcontainer, and that directory is only on the *terminal's* `PATH`, so `Sys.which("air")` comes up empty in an R console; the check looks in the extension directory too, and warns rather than failing if it finds nothing.

By default only rows with an evaluated status are checked, matching what the harmoniser actually runs.
Pass `active_only = FALSE` to include `in_progress` and other skipped rows while you are still drafting them.

### Fixing formatting

`fix = TRUE` writes each `suggestion` back into the `expression` column and reports what it changed in the `fixed` column:

```r
check_harmonisation_exprs(fix = TRUE)
```

Only `unformatted` rows are rewritten; the other three issues need a human decision and are left untouched.
Every other column is read and written as text, so a fix cannot disturb cells it was not asked to change.

Expressions are formatted to their own settings rather than the ones in the repository's `air.toml`, which continues to govern the `.R` files:

| Setting | Default | Why |
|---------|---------|-----|
| `line_width` | `120` | An expression in a CSV cell has no surrounding indentation to pay for, and the 80-column default explodes ordinary recodes into one argument per line. |
| `persistent_line_breaks` | `FALSE` | Air normally keeps a call expanded if the source has a line break straight after the opening `(`. In a CSV cell that break reflects hand-editing rather than intent, so width alone should decide. |

Both are arguments, so you can widen or narrow a run without editing any configuration.

## Testing your harmonisation

```r
targets::tar_make(names = harmonised_BPIPD_9999) # your dataset id
targets::tar_read(harmonised_BPIPD_9999)
```

Check derived values look sensible (tabulate `sex`, summarise `age_years`, …).
Then run the repository test suite, which validates all harmonisation metadata without needing the data:

```r
testthat::test_dir("tests/testthat")
```

Both run on every pull request — CI validates the metadata and expressions, but only your local `tar_make()` sees real values.
