---
title: "Harmonising variables"
---

Harmonisation maps a dataset's [tidied table](tidy-scripts.qmd) onto the project [dataschema](../reference/dataschema.qmd), so that every study contributes the same variables with the same meanings.
It is entirely table-driven: you edit CSV files, not R scripts.

The files involved:

| File | Purpose |
|------|---------|
| `harmonisation/dataschema.csv` | The shared schema — one row per harmonised variable (shown in the [Dataschema reference](../reference/dataschema.qmd)) |
| `harmonisation/datasets/BPIPD-<id>/variables.csv` | How *this* dataset derives each schema variable |
| `harmonisation/lookups/*.csv` | Shared recode tables reused across studies (e.g. ethnicity) |
| `harmonisation/datasets/BPIPD-<id>/lookups/*.csv` | Large one-off recode tables specific to this dataset |

## variables.csv

One row per dataschema variable, in dataschema order.
The template copy starts with every variable set to `in_progress`; harmonising a dataset means working through the rows.

| Column | Meaning |
|--------|---------|
| `target_variable` | The dataschema variable being derived. Every non-system schema variable must have a row (`dataset_id` and `dataset_name` are injected by the pipeline). |
| `status` | How well the study supports this variable (table below). |
| `source_columns` | Which tidied columns the mapping uses — documentation for reviewers, not executed. |
| `expression` | R code that computes the value (required unless the status says the variable can't be derived). |
| `notes` | Anything a reviewer needs: assumptions, DOIs, caveats. |
| `lookup_table` | Basename of a lookup CSV this expression uses, if any (e.g. `ethnicity` for `harmonisation/lookups/ethnicity.csv`). |

A few illustrative rows (the column names are fictional, but the shapes mirror real mappings in the project):

```csv
target_variable,status,source_columns,expression,notes,lookup_table
participant_id,compatible,serial,as.character(serial),Assume serial is the participant ID,
wave,unavailable,,,Cross-sectional study,
data_year,inferred,,2016,Not in the data; taken from the study protocol,
sex,compatible,sex_raw,"recode_values(haven::as_factor(sex_raw), ""F"" ~ ""Female"", ""M"" ~ ""Male"")",Convert labels to harmonised values,
```

## Statuses

Statuses follow the [Maelstrom Research](https://www.maelstrom-research.org/) harmonisation vocabulary:

```{r}
#| label: status-table
statuses <- tibble::tribble(
  ~Status, ~Meaning, ~Evaluated,
  "compatible",
  paste(
    "The study variable can generate the target with an explicit,",
    "scientifically defensible transformation and without meaningful",
    "loss of meaning."
  ), "Yes",
  "partial",
  paste(
    "The study only supports part of the target construct, or the",
    "harmonised value requires a simplification that should be",
    "interpreted with caution."
  ), "Yes",
  "proximate",
  paste(
    "A categorical subtype of partial: the study captures a close but",
    "not fully equivalent category structure."
  ), "Yes",
  "inferred",
  paste(
    "The value is derived from study metadata or surrounding context",
    "(e.g. a fixed collection year from the study's publication)",
    "rather than a like-for-like study variable."
  ), "Yes",
  "incompatible",
  paste(
    "Related information exists, but it cannot be transformed into a",
    "defensible version of the target variable."
  ), "No — output is NA",
  "unavailable",
  "The study does not collect the required information.",
  "No — output is NA",
  "in_progress",
  "The mapping has not been completed yet (the template default).",
  "No — output is NA"
)
knitr::kable(statuses)
```

```{r}
#| label: status-drift-check
#| include: false
# Fail the site build if this table drifts from the pipeline's status set.
source("../R/config.R")
stopifnot(setequal(statuses$Status, bp_harmonisation_status_values()))
```

Rows with an "evaluated" status must have an `expression`; rows with the other three statuses are skipped and the variable is filled with typed `NA` values for that dataset.

## Writing expressions

An expression is a single R expression evaluated against the tidied table.
Its result is recycled to the table's row count (so a constant like `2016` is fine) and cast to the schema's declared type.
Inside an expression you can reference:

- **columns of the tidied table** directly by name (`serial`, `sex_raw`);
- `tbl` or `analysis_base` — the whole tidied table, if you need it;
- `spec` — the parsed `dataset.yaml` (e.g. `spec$dataset_id`);
- `lookup_<name>` — any lookup table named in the row's `lookup_table` column (e.g. `lookup_ethnicity`).

Expressions run in a **sandboxed environment**, not your R session.
Bare function calls resolve from base R plus this curated allowlist:

```{r}
#| label: allowlist-table
source("../R/helpers.R")
source("../R/harmonise/harmonisation_functions.R")
fns <- bp_harmonisation_functions()
origin <- vapply(
  fns,
  function(fn) {
    env_name <- environmentName(topenv(environment(fn)))
    if (env_name %in% c("R_GlobalEnv", "")) "pipeline helper" else env_name
  },
  character(1)
)
allowlist <- tibble::tibble(
  Function = paste0("`", names(fns), "()`"),
  From = origin
)
knitr::kable(allowlist)
```

Namespaced calls such as `haven::as_factor()` also work for anything not on the list.
If an expression calls a bare function that is neither base R nor allowlisted, the test suite fails with a message telling you to either namespace it or add it to `bp_harmonisation_functions()` (`R/harmonise/harmonisation_functions.R`) — extending the allowlist is a one-line change, reviewed like any other.

Typical recodes use `recode_values()` with formula syntax:

```r
recode_values(
  haven::as_factor(sex_raw),
  "F" ~ "Female",
  "M" ~ "Male"
)
```

## Lookup tables

Three tiers, from smallest to largest:

1. **Inline** — short, study-specific recodes stay in the expression itself (like the `sex` example above).
2. **Shared lookups** (`harmonisation/lookups/*.csv`) — recurring recodes that should be reviewed centrally, such as ethnicity or maternal education.
   These must have the columns `dataset_id`, `target_variable`, `source_value`, `harmonised_value`, `mapping_status`, `notes`, with one row per source value per dataset.
3. **Dataset lookups** (`harmonisation/datasets/BPIPD-<id>/lookups/*.csv`) — mappings too large to read inline but not shared across studies.
   At minimum they need `source_value` and `harmonised_value` columns.

To use one, put its basename in the row's `lookup_table` column and call `lookup_values()` with the matching `lookup_<name>` object.
For example, an `ethnicity` row with `lookup_table` = `ethnicity`:

```r
lookup_values(
  haven::as_factor(ethnicity_raw),
  lookup_ethnicity,
  dataset_id = spec$dataset_id,
  target_variable = "ethnicity"
)
```

The `dataset_id`/`target_variable` arguments filter a shared lookup down to this dataset's rows; omit them for a dataset-specific lookup.
A lookup name must exist in exactly one of the two locations — the validator errors if it is missing from both or present in both.

## Keeping variables.csv in sync with the dataschema

When the dataschema gains a variable, every dataset's `variables.csv` needs a matching row.
Don't do this by hand:

```r
targets::tar_source()
sync_harmonisation_vars() # all datasets
```

This appends any missing variables as `in_progress` and reorders rows to dataschema order.
The validator (and therefore CI) rejects a `variables.csv` with missing, duplicate, or unknown target variables.

## Testing your harmonisation

```r
targets::tar_make(names = harmonised_BPIPD_9999) # your dataset id
targets::tar_read(harmonised_BPIPD_9999)
```

Check derived values look sensible (tabulate `sex`, summarise `age_years`, …).
Then run the repository test suite, which validates all harmonisation metadata without needing the data:

```r
testthat::test_dir("tests/testthat")
```

Both run on every pull request — CI validates the metadata and expressions, but only your local `tar_make()` sees real values.
