---
title: "Harmonisation progress"
---

Generated from the specs and mapping files in the repository on every site build.
A dataset appears here once its folder under `harmonisation/datasets/` contains a `dataset.yaml`.

```{r}
#| label: load-metadata
#| include: false
datasets_dir <- "../harmonisation/datasets"
spec_files <- fs::dir_ls(datasets_dir, recurse = 1, glob = "*dataset.yaml")

specs <- lapply(spec_files, yaml::read_yaml)
dataset_dirs <- fs::path_dir(spec_files)

inventory <- tibble::tibble(
  dataset_id = vapply(specs, function(s) s$dataset_id, character(1)),
  status = vapply(specs, function(s) s$status, character(1)),
  waves = vapply(
    specs,
    function(s) if (is.null(s$waves)) 1L else length(s$waves),
    integer(1)
  ),
  has_mappings = fs::file_exists(fs::path(dataset_dirs, "variables.csv"))
)

read_mappings <- function(dir, dataset_id) {
  path <- fs::path(dir, "variables.csv")
  if (!fs::file_exists(path)) {
    return(NULL)
  }
  vars <- readr::read_csv(path, show_col_types = FALSE)
  tibble::tibble(
    dataset = paste0("BPIPD-", dataset_id),
    target_variable = vars$target_variable,
    status = vars$status
  )
}

mappings <- dplyr::bind_rows(
  mapply(
    read_mappings,
    dataset_dirs,
    inventory$dataset_id,
    SIMPLIFY = FALSE
  )
)
```

## Datasets

```{r}
#| label: dataset-inventory
inventory |>
  dplyr::transmute(
    Dataset = paste0("BPIPD-", dataset_id),
    Status = status,
    Waves = waves,
    `Variables mapped` = ifelse(has_mappings, "Yes", "Not yet")
  ) |>
  knitr::kable()
```

`planned` and `blocked` datasets are excluded from the pipeline; `in_progress` and `harmonised` datasets are tidied and (once mappings exist) harmonised into the combined analysis dataset.

## Mapping status by dataset

How each dataset's variable mappings break down across the [harmonisation statuses](../guides/harmonisation.qmd#statuses):

```{r}
#| label: status-summary
status_order <- c(
  "compatible", "partial", "proximate", "inferred",
  "incompatible", "unavailable", "in_progress"
)
mappings |>
  dplyr::count(dataset, status) |>
  dplyr::mutate(status = factor(status, levels = status_order)) |>
  tidyr::pivot_wider(
    names_from = status,
    values_from = n,
    values_fill = 0L,
    names_expand = TRUE
  ) |>
  dplyr::rename(Dataset = dataset) |>
  knitr::kable()
```

## Variable-by-dataset matrix

The full picture: one row per dataschema variable, one column per dataset, showing the mapping status.
Use the filters to focus on a domain's variables or find gaps.

```{r}
#| label: status-matrix
#| column: page
dataschema <- readr::read_csv(
  "../harmonisation/dataschema.csv",
  show_col_types = FALSE
)

matrix_tbl <- mappings |>
  tidyr::pivot_wider(
    names_from = dataset,
    values_from = status
  )

matrix_tbl <- dataschema |>
  dplyr::transmute(target_variable = variable_name, domain) |>
  dplyr::inner_join(matrix_tbl, by = "target_variable") |>
  dplyr::rename(variable = target_variable)

dataset_cols <- setdiff(names(matrix_tbl), c("variable", "domain"))

status_colours <- c(
  compatible = "#c7e9c0",
  partial = "#fee391",
  proximate = "#fee391",
  inferred = "#c6dbef",
  incompatible = "#fcbba1",
  unavailable = "#f0f0f0",
  in_progress = "#d9d9d9"
)

table <- DT::datatable(
  matrix_tbl,
  filter = "top",
  rownames = FALSE,
  options = list(pageLength = 25, scrollX = TRUE)
)
DT::formatStyle(
  table,
  columns = dataset_cols,
  backgroundColor = DT::styleEqual(
    names(status_colours),
    unname(status_colours)
  )
)
```
